---
title: "bigger_n_better"
author: "SarahKinz"
date: "5/28/2021"
output: html_document
---
What do I want to know about each spring... Let's stick to basics

How many springs are present in the state?
How many springs have discharge?
How many springs have nitrate?
How many are within range range?
How many are in range, but do not have the other measurement?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dataRetrieval)
library(tidyverse)
library(dplyr)
library(naniar)
library(ggplot2)
```

I got the bounding boxes from https://anthonylouisdagostino.com/bounding-boxes-for-all-us-states/.

### Creating BBox that you did not actually need
```{r}
VA <- c(-83.675395, 36.540738, -75.242266,	39.466012)
WV<- c(-83.675395, 36.540738, -75.242266, 39.466012)
MD<- c(-79.487651, 37.911717, -75.048939, 39.723043)
DE<- c(-75.788658, 38.451013, -75.048939, 39.839007)
NY<- c(-79.762152, 40.496103, -71.856214, 45.01585)
PA<- c(-80.519891, 39.7198, -74.689516, 42.26986)
DC <- c(-77.119759, 38.791645, -76.909395, 38.99511)

```

### VIRGINIA
```{r}

### I am reading in the data to see what I have

### VIRGINIA

VA_sites <- whatNWISdata(stateCd = "VA") 

VA_sites <- VA_sites%>%filter(site_tp_cd == "SP")

#255 unique sites
 
siteNumber <- unique(VA_sites$site_no)
startDate <- ""
endDate <- ""

#Discharge

parameterCd<- "00060"

VA_discharge <- readNWISdv(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 4 sites

VA_discharge_table<-VA_discharge%>%mutate(Observation= 1, GPM=Flow*448.83)%>%group_by(site_no)%>%summarise(Median= median(GPM), Min= min(GPM), Max= max(GPM), n= sum(Observation))

more_VA_discharge<- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

more_VA_discharge_table<-more_VA_discharge%>%mutate(Observation= 1, GPM=result_va*448.83)%>%group_by(site_no)%>%summarise(Median= median(GPM), Min= min(GPM), Max= max(GPM), n= sum(Observation))


double_VA_discharge<- rbind(VA_discharge_table, more_VA_discharge_table)%>%filter(Median> 90, Median<1100)

## 11 sites

#Nitrate

#Ask about using 00631 vs 00630, I am using 00630 for now because it matches

parameterCd <- ("00630")

VA_nitrate_unfiltered <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

VA_nitrate_table_unfiltered<-VA_nitrate_unfiltered%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))

parameterCd <- ("00631")

VA_nitrate_filtered <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

VA_nitrate_table_filtered<-VA_nitrate_filtered%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))


VA_combo_nirate_table<- rbind(VA_nitrate_table_unfiltered, VA_nitrate_table_filtered)

## 8 sites


# Temp

parameterCd<- "00010"

VA_temp <- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 198 sites

#pH

parameterCd<- c("00400", "00403")

VA_pH <-  readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 211 sites 

#Dissolved Oxygen

#Dissolved Organic Carbon

# you can add the state data you have here:)

# Joining VA nitrate together... Just reading it in

DEQ_nitrate<- read_csv("DEQ_Water_Quality.csv")


#Joining VA discharge together... Just reading it in

DEQ_discharge<- read_csv("VDEQ_Springs.csv")

# JOining VA discharge and nitrate together... Probs not yet. 


```


### VIRGINIA ANALYSIS (trying to do remember what you lost)
```{r}

# consolidating DEQ_nitrate

refined_DEQ_WQ<- DEQ_nitrate%>%select(OBJECTID, ORIGINAL_WQ_DATA_SET, USGS_SITE_NUMBER, DEQ_WELL_NUMBER, OWNER, LAT83dd, LONG83dd, SAMPLE_COLLECTION_DATE, pH_FIELD_STANDARD_UNITS, pH_STANDARD_UNITS, CARBON_TOTAL_mg_L, CARBON_INORGANIC_FILTERED_mg_L, CARBON_INORGANIC_mg_L, CARBON_ORGANIC_FILTERED_mg_L, CARBON_ORGANIC_mg_L, OWNER_SITE,NITRATE_PLUS_NITRITE_FILTERED_mg_L_N,NITRATE_PLUS_NITRITE_mg_L_N, NITRATE_FILTERED_mg_L_N, NITRATE_mg_L_N, SPRING_NAME, SITE_COMMENTS, TOPO_SETTING, DATA_COLLECTOR_SITE, GENERAL_LITHOLOGY, MAJOR_BASIN_NAME, VDOF_LAND_USE, PROVINCE)

# smaller nitrate

small_DEQ_nitrate<- DEQ_nitrate%>%select(DEQ_WELL_NUMBER, NITRATE_PLUS_NITRITE_mg_L_N)

# consolidating DEQ_discharge data

refined_DEQ_discharge<- DEQ_discharge%>% select(OBJECTID, DEQ_NUMBER, LATNAD83DD, LONGNAD83D, DATE_MEASURED, DISCHARGE_METHOD, DISCHARGE_GPM, DATA_SOURCE, pH, FIELD_COMMENTS, SPRING_NAME, USGS_SITE_ID, TOPO_SETTING, COUNTY)

#smaller discharge

small_DEQ_discharge<- DEQ_discharge%>% select(DEQ_NUMBER, DISCHARGE_GPM)

# joined summary of springs in range

small_DEQ_nitrate_table<-small_DEQ_nitrate%>%filter(NITRATE_PLUS_NITRITE_mg_L_N>=-1)%>%mutate(Observation=1)%>%
group_by(DEQ_WELL_NUMBER)%>%summarise('Median Nitrate (mg/L)'=median(NITRATE_PLUS_NITRITE_mg_L_N), 'Mean Nitrate (mg/L)'= mean(NITRATE_PLUS_NITRITE_mg_L_N), 'Max Nitrate (mg/L)'= max(NITRATE_PLUS_NITRITE_mg_L_N), 'Min nitrate (mg/L)'=min(NITRATE_PLUS_NITRITE_mg_L_N), Observation= sum(Observation))

nitrate_criteria<- small_DEQ_nitrate_table%>%filter(`Median Nitrate (mg/L)`>= 3.0)


####### THERE ARE 49 SPRINGS THAT MEET NITRATE CRITERIA #######

small_DEQ_discharge_table<-small_DEQ_discharge%>%filter(DISCHARGE_GPM>=-1)%>%mutate(Observation=1)%>%group_by(DEQ_NUMBER)%>%summarise('Median Discharge (GPM)'=median(DISCHARGE_GPM), 'Mean Discharge (GPM)'= mean(DISCHARGE_GPM), 'Max Discharge (GPM)'= max(DISCHARGE_GPM), 'Min Discharge (GPM)'=min(DISCHARGE_GPM), Observation= sum(Observation))


discharge_criteria<-small_DEQ_discharge_table%>%filter(`Median Discharge (GPM)`<= 1100 & `Median Discharge (GPM)`>= 90)

####### THERE ARE 133 SPRINGS THAT MEET DISCHARGE CRITERIA #######

small_picture<- full_join(small_DEQ_discharge_table, small_DEQ_nitrate_table, c("DEQ_NUMBER"="DEQ_WELL_NUMBER"))

smaller_picture<-small_picture%>%filter(`Median Discharge (GPM)`<= 1100,`Median Discharge (GPM)`>= 90, `Median Nitrate (mg/L)`>= 3.00)

####### THERE ARE 8 SPRINGS THAT MEET BOTH CRITERIA #######

### Finding out the info to create the bargraph 

graph_info<- full_join(small_DEQ_discharge, small_DEQ_nitrate, c("DEQ_NUMBER"="DEQ_WELL_NUMBER"))

graph_info[graph_info == -9999] <- NA

graph_info<-graph_info%>%group_by(DEQ_NUMBER)%>%summarise('Median Nitrate (mg/L)'=median(NITRATE_PLUS_NITRITE_mg_L_N, na.rm=TRUE), 'Median Discharge (GPM)'= median(DISCHARGE_GPM, na.rm=TRUE))

test<- graph_info%>%
  mutate(category = case_when(`Median Discharge (GPM)`== as.character(NA) &
                              `Median Nitrate (mg/L)`== as.character(NA) ~ "no data",
                              `Median Discharge (GPM)`== as.character(NA) &
                              `Median Nitrate (mg/L)`<= 0 ~ "no discharge data",
                              `Median Discharge (GPM)`<= 0 &
                              `Median Nitrate (mg/L)`== as.character(NA) ~ "no nitrate data",
                              `Median Discharge (GPM)`<= 1100  &
                              `Median Discharge (GPM)`>= 90  &
                              `Median Nitrate (mg/L)`>= 3.0 ~ "in range",
                              `Median Discharge (GPM)`<= 90  &
                              `Median Discharge (GPM)`>= 1100  &
                              `Median Nitrate (mg/L)`<= 3.0 ~ "out of range"))
                              
    

### Actually creating that stupid graph again

graph<-data.frame("Category"= c("No Nitrate Data", "No Discharge Data",
          "No Data","Out of Range", "In Range"),
           "Quantity"= c(232,117, 157, 108, 8))

ggplot(data=graph, aes(x=Category, y=Quantity)) +
  geom_bar(stat="identity", color="blue", fill= "blue")+
  theme_classic()

```





### WEST VIRGINIA
```{r}
### West Virginia

#Getting data from USGS

WV_sites <- whatNWISdata(stateCd = "WV") 

WV_sites <- WV_sites%>%filter(site_tp_cd == "SP")

siteNumber <- unique(WV_sites$site_no)
startDate <- ""
endDate <- ""

#Discharge

parameterCd<- "00060"

WV_discharge <- readNWISdv(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 3 sites

WV_discharge_table<-WV_discharge%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(Flow), Min= min(Flow), Max= max(Flow), n= sum(Observation))

more_WV_discharge<- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 2 sites

more_WV_discharge_table<-more_WV_discharge%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))

# Temp

parameterCd<- "00010"

WV_temp <- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 190 sites

#Nitrate

parameterCd <- ("00631")

WV_nitrate <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 92 sites

WV_nitrate_table<-WV_nitrate%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))

#pH

parameterCd<- c("00400", "00403")

WV_pH <-  readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 188 sites

#Dissolved Oxygen

#Dissolved Organic Carbon


```


### MARYLAND
```{r}
## Getting Data from USGS
MD_sites <- whatNWISdata(stateCd = "MD") 

MD_sites <-MD_sites%>%filter(site_tp_cd == "SP")


siteNumber <- unique(MD_sites$site_no)
startDate <- ""
endDate <- ""

#Discharge

parameterCd<- "00060"

MD_discharge <- readNWISdv(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

MD_discharge_table<-MD_discharge%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(Flow), Min= min(Flow), Max= max(Flow), n= sum(Observation))

## 1 site

more_MD_discharge<- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 0 sites 

# Temp

parameterCd<- "00010"

MD_temp <- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 127 sites

#Nitrate

parameterCd <- ("00631")

MD_nitrate <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 61 sites

MD_nitrate_table<-MD_nitrate%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))

#pH

parameterCd<- c("00400", "00403")

MD_pH <-  readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)


## 142 sites

#Dissolved Oxygen

#Dissolved Organic Carbon

```

Delaware seems to be accurate. There are two other inactive springs in Delaware, but there is no data for them aside from presences. The one site that has data does not have the data I need.

### DELAWARE
```{r}
## Getting Data from USGS
DE_sites <- whatNWISdata(stateCd = "DE") 

DE_sites <-DE_sites%>%filter(site_tp_cd == "SP")

siteNumber <- unique(DE_sites$site_no)
startDate <- ""
endDate <- ""

#Discharge

parameterCd<- "00060"

DE_discharge <- readNWISdv(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 0 sites

more_DE_discharge<- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 0 sites

# Temp

parameterCd<- "00010"

DE_temp <- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 0 sites

#Nitrate

parameterCd <- ("00631")

DE_nitrate <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 0 sites

#pH

parameterCd<- c("00400", "00403")

DE_pH <-  readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)


## 0 sites

#Dissolved Oxygen

#Dissolved Organic Carbon

```

### NEW YORK
```{r}
## Getting Data from USGS
NY_sites <- whatNWISdata(stateCd = "NY") 

NY_sites <-NY_sites%>%filter(site_tp_cd == "SP")

siteNumber <- unique(NY_sites$site_no)
startDate <- ""
endDate <- ""

#Discharge

parameterCd<- "00060"

NY_discharge <- readNWISdv(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 0 sites

more_NY_discharge<- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 0 sites


# Temp

parameterCd<- "00010"

NY_temp <- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 96 sites

#Nitrate

parameterCd <- ("00631")

NY_nitrate <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 52 sites

NY_nitrate_table<-NY_nitrate%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))

#pH

parameterCd<- c("00400", "00403")

NY_pH <-  readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 150 sites


#Dissolved Oxygen

#Dissolved Organic Carbon

```


### PENNSLVANYIA
```{r}
## Getting Data from USGS
PA_sites <- whatNWISdata(stateCd = "PA") 

PA_sites <-PA_sites%>%filter(site_tp_cd == "SP")


siteNumber <- unique(PA_sites$site_no)
startDate <- ""
endDate <- ""

#Discharge

parameterCd<- "00060"

PA_discharge <- readNWISdv(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 1 site

PA_discharge_table<-PA_discharge%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(Flow), Min= min(Flow), Max= max(Flow), n= sum(Observation))

more_PA_discharge<- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 59 sites

more_PA_discharge_table<-more_PA_discharge%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))

# Temp

parameterCd<- "00010"

PA_temp <- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 283 sites

#Nitrate

parameterCd <- ("00631")

PA_nitrate <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 80 sites

PA_nitrate_table<-PA_nitrate%>%mutate(Observation= 1)%>%group_by(site_no)%>%summarise(Mean= mean(result_va), Min= min(result_va), Max= max(result_va), n= sum(Observation))

#pH

parameterCd<- c("00400", "00403")

PA_pH <-  readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 422 sites


#Dissolved Oxygen

#Dissolved Organic Carbon


## Using data from the state agency

PA_discharge<- read.csv("PA_spring_Discharge.csv")
PA_discharge<- PA_discharge%>%filter(Discharge.gpm.>=90)%>%
  filter(Discharge.gpm.<=1100)

PA_chemical<- read.csv("PA_Springs_Chemical.csv")

PA_chemical_small<- PA_chemical%>%
  pivot_wider(names_from = ParamLongName, values_from = QWValue)%>%select(PASpringID,QWDescription,County,Date,`ELEVATION (FEET NGVD) `,`SAMPLING DEPTH (FEET) `,`GAGE HEIGHT FEET `,`PH WATER WHOLE LABORATORY STANDARD UNITS `,`PH WATER WHOLE FIELD STANDARD UNITS `,`FLOW RATE INSTANTANEOUS (GALLONS/MINUTE) `,`CARBON DIOXIDE DISSOLVED (MG/L AS CO2) `,`CARBON ORGANIC TOTAL (MG/L AS C) `,`OXYGEN DISSOLVED (MG/L) `,`DISCHARGE INSTANTANEOUS CUBIC FEET PER SECOND `,`DISCHARGE CUBIC FEET PER SECOND `,`NITROGEN DISSOLVED (MG/L AS N) `,`NITROGEN NITRITE DISSOLVED (MG/L AS NO2) `,`NITROGEN NITRITE PLUS NITRATE DISSOLVED (MG/L AS N) `,`NITROGEN NITRITE DISSOLVED MG/L AS N `,`NITROGEN NITRATE TOTAL (MG/L AS N) `,`NITROGEN TOTAL (MG/L AS N) `,`NITROGEN NITRATE TOTAL (MG/L AS N) `,`NITROGEN NITRATE DISSOLVED (MG/L AS NO3) `,`NITROGEN NITRATE DISSOLVED (MG/L AS N) `,`DEPTH BELOW LAND SURFACE (WATER LEVEL) (FEET) `,`WATER TEMPERATURE DEGREES CELSIUS `)

# There are some hecking high`NITROGEN NITRITE PLUS NITRATE DISSOLVED (MG/L AS N) `) values.


```



### DISTRICT OF COLUMBIA
```{r}
## Getting Data from USGS
DC_sites <- whatNWISdata(stateCd = "DC") 

DC_sites <-DC_sites%>%filter(site_tp_cd == "SP")

siteNumber <- unique(DC_sites$site_no)
startDate <- ""
endDate <- ""

#Discharge

parameterCd<- "00060"

DC_discharge <- readNWISdv(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 0 sites

more_DC_discharge<- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()
# Temp

parameterCd<- "00010"

DC_temp <- readNWISqw(siteNumber, parameterCd, 
        startDate, endDate)%>% renameNWISColumns()

## 1 site

#Nitrate

parameterCd <- ("00631")

DC_nitrate <- readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

## 1 site

#pH

parameterCd<- c("00400", "00403")

DC_pH <-  readNWISqw(siteNumber, parameterCd, 
      startDate, endDate)

#Dissolved Oxygen

#Dissolved Organic Carbon


```

Doing it the old fashioned way

# APPROXIMATELY THE ENTIRE CHESAPEAKE BAY WATERSHED 
```{r}

### Standard Detail Data (coordinates, site number, etc)

inactive_p1<- read_csv("Inactive Springs Bradford to Rochester/NWISMapperExport.csv")

inactive_p2<- read_csv("Inactive Springs Charlottesville to Martinsburg/NWISMapperExport.csv")

inactive_p3<- read_csv("Inactive Springs Danville to Charlottesville/NWISMapperExport.csv")

inactive_p4<- read_csv("Inactive Springs Martinsburg to State College/NWISMapperExport.csv")

inactive_p5<- read_csv("Inactive Springs State College to Bradford/NWISMapperExport.csv")

active_p1<- read_csv("Active Springs Charlottesville to Martinsburg/NWISMapperExport.csv")

active_p2<- read_csv("Active Springs Bradford to Rochester/NWISMapperExport.csv")

active_p3<- read_csv("Active Spring Martinsburg to State College/NWISMapperExport.csv")

#### Water Quality Data

wq_inactive_p1<- read_csv("Inactive Springs WQ Bradford to Rochester/NWISMapperExport.csv")

wq_inactive_p2<- read_csv("Inactive Springs WQ Charlottesville to Martinsburg/NWISMapperExport.csv")

wq_inactive_p3<- read_csv("Inactive Springs WQ Danville to Charlottesville/NWISMapperExport.csv")

wq_inactive_p4<- read_csv("Inactive Springs WQ Martinsburg to State College/NWISMapperExport.csv")

wq_inactive_p5<- read_csv("Inactive Springs WQ State College to Bradford/NWISMapperExport.csv")

wq_active_p1<- read_csv("Active Springs WQ Charlottesville to Martinsburg/NWISMapperExport.csv")

wq_active_p2<- read_csv("Active Springs WQ Bradford to Rochester/NWISMapperExport.csv")

wq_active_p3<- read_csv("Active Spring WQ Martinsburg to State College/NWISMapperExport.csv")

```

